---
title: "Class Code for 1021"
author: "Maxwell Cornellier"
date: "10/21/2020"
output: 
  html_document:  
    number_sections: true
    toc: true
    toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

## Import Data

First thing we will do is to **import data**

```{r, cache=TRUE}
# Using read.csv for reading the csv file
nyc <- read.csv("nycflights.csv")

# Now we will explore this data
str(nyc)
head(nyc)
tail(nyc)

```

## Clean Data

We Will now need to clean this data.

```{r}
# Delete the column named "X"
nyc$X <- NULL

# Recode columns as factors as needed
nyc$carrier <- as.factor(nyc$carrier)
nyc$flight <- as.factor(nyc$flight)
nyc$tailnum <- as.factor(nyc$tailnum)
nyc$origin <- as.factor(nyc$origin)
nyc$dest <- as.factor(nyc$dest)

# Re-check summary and str

str(nyc)
summary(nyc)
```

# Calculate Intersting Metrics 

```{r}
#Average departure delay
mean(nyc$dep_delay, na.rm = TRUE)
#Average air-time
mean(nyc$air_time, na.rm = TRUE)
# Standard Deviation
sd(nyc$air_time, na.rm = TRUE)

# Which origin airport has the most flights
```

# Data Exploration

## Slice and Dice Data - Filter

> Logical Indexing 


```{r}
nyc[ 3 , 5 ]

nyc[2:10,3]

nyc[1453:1525,"arr_delay"]

nyc[1453:1525,c("dep_delay" , "arr_delay")]

```

## Complicated Logical Indexing 

## How many flights from LGA to DTW in 2013 

```{r}
lga2dtw <- nyc[nyc$origin == "LGA" & nyc$dest == "DTW" ,]
nrow(lga2dtw)
```

### Random Sampling 

```{r}
nyc[sample(1:nrow(nyc),5), ]
```

### Example: All flights delayed by more than 60 min

```{r}
nyc60min <- nyc[nyc$dep_delay >= 60 & nyc$carrier == "DL", ]
nrow(nyc60min)
```

### Example 2: All flights more than 100 miles 

```{r}
nyc1000miles<-nyc[nyc$distance >= 1000,]
nrow(nyc1000miles)
```

## Sorting Data

```{r}
head(nyc[order(nyc$arr_delay, decreasing = T), ])
```

# Average Arrival Delay 
```{r}
# Use function tapply 

arr_delay_by_airline <- round(tapply(nyc$arr_delay, nyc$carrier,mean,na.rm = T),2)

barplot(arr_delay_by_airline)
```

## Sorted barplot

```{r}

barplot(sort(arr_delay_by_airline, decreasing = T))
```

## Highest Average Arrival Delay 

```{r}

max_delay_airlines <- names(arr_delay_by_airline[arr_delay_by_airline == max(arr_delay_by_airline)])


```



## Inline R Code 

The airline with the highest average arrival delay is: `r max_delay_airlines`


